Automated Design of Nanophotonic Waveguide Couplers

### 2011-10-07 presentations/ob1_wg_paper/paper.tex
First draft done.

\documentclass[letterpaper,10pt]{article}
\usepackage{graphicx}
% \usepackage{osameet2}
\usepackage{amsmath,amssymb}

\begin{document}
\title{Automated Design of Nanophotonic Waveguide Couplers}
\author{Jesse Lu and Jelena Vu\v{c}kovi\'{c}}
% \address{Stanford University, Stanford, California, USA.}
% \email{jesselu@stanford.edu}

\maketitle
\begin{abstract}
We demonstrate a design algorithm which automatically generates 
    wavelength-scale coupling devices between arbitrary waveguide modes 
    with high efficiency.
Our algorithm is computationally fast, 
    can be extended to multiple dimensions,
    and requires no trial-and-error.
\end{abstract}
% \ocis{230.7370, 130.3990.}

@@

% Why is waveguide mode conversion important?
\subsection{The importance of waveguide mode conversion}
Optical mode conversion, 
    the efficient transfer of photons from one guided mode to another,
    is a fundamental requirement in nanophotonics.
Efficient conversion between waveguides modes
    is critical in the many cases, including:
\begin{enumerate}
    \item Coupling to and from optical fiber\cite{}, 
        to communicate with the outside world.
    \item Coupling between various nanophotonic waveguides, 
        since different waveguides are best suited for different applications.
        For example, ridge waveguides seem ideal for low-loss transport\cite{},
            but other waveguides, 
            such as photonic crystal waveguides or slot waveguides,
            may be better suited for slow-light\cite{} 
            or energy-focusing devices\cite{}.
    \item Coupling between different materials.
       This is to couple between passive, active\cite{}, 
        and non-linear\cite{} materials. 
\end{enumerate}

% Why are current solutions insufficient?
\subsection{Common approaches to designing waveguide couplers}
% Brute force
Brute-force parameter search is the most popular nanophotonic design strategy 
    to-date because of its sheer simplicity\cite{}.
Although it may be suitable for tuning existing designs\cite{},
    the parameter space for most practical devices is simply too large
    for such a strategy to be tractable.

% Adiabatic mode conversion (large devices, symmetry breaking)
Adiabtic mode conversion strategies have been succesful
    for certain fiber-waveguide\cite{} and waveguide-waveguide\cite{} couplers,
    although resulting devices are quite large.
However, adiabatic strategy cannot be used in many important cases such as
    coupling from ridge to some photonic crystal waveguides,
    coupling in the out-of-plane direction, and
    coupling between modes of opposite symmetry.

% Local optimization
Optimization methods based on local derivatives seem very promising\cite{},
    in that they are both much faster than brute-force methods and
    more adaptable than adiabatic strategies.
However, these methods still require that every updated design be simulated
    at least once,
    and for the user to supply an initial design.

% How does your algorithm solve this problem?
\subsection{Advantages of objective-first design}
We present an ``objective-first'' approach to nanophotonic design, 
    and apply it to the problem of high-efficiency waveguide couplers.
The resulting algorithm
\begin{itemize}
    \item does not employ brute-force parameter searches,
    \item does not require a good initial design,
    \item is computationally fast (no simulations required),
    \item generates couplers between seemingly arbitrary waveguide modes, and
    \item can generate these couplers in a very small footprint. 
\end{itemize}

@@

The differences between Eqs.~\ref{eq:adj} and \ref{eq:ob1} are
\begin{enumerate}
    \item in Eq.~\ref{eq:adj}, we attempt to decrease the design objective 
            (Eq.~\ref{eq:adj:obj}), 
        while in Eq.~\ref{eq:ob1}, the design objective is kept at zero 
            (Eq.~\ref{eq:ob1:con}); 
    \item in Eq.~\ref{eq:adj}, we always satisfy the underlying physics 
            (Eq.~\ref{eq:adj:con}),
        while in Eq.~\ref{eq:ob1}, physics is not satisfied,
            since the physics residual is generally non-zero
            (Eq.~\ref{eq:ob1:obj}).
\end{enumerate}

@@

The first practical implication of the objective-first approach 
    is that the number of independent variables is increased to include
    both $x$ and $p$.
In Eq.~\ref{eq:adj}, the constraint that physics must be satisfied, $g(x,p)=0$, 
    essentially forces $x$ to be dependent on $p$,
    since the choice of $p$ implicitly determines the value of $x$
    (there is generally a one-to-one mapping from $p$ to $x$).
In contrast, Eq.~\ref{eq:ob1} allows both $x$ and $p$ to vary independently,
    because the constraint, $f(x)=0$, is only a function of $x$.

@@

Together, these implications result in a method
    that is computationally fast (since it does not require simulation), and
    that can be applied to non-intuitive problems 
    where a functional starting design is not readily available.

% How is it applied to waveguide coupler design?
\subsection{Objective-first approach to waveguide coupler design}
We apply an objective-first approach to the problem of designing 
    two-dimensional nanophotonic waveguide couplers.

@@

This problem contains many local minima (it is non-convex\cite{});
    however, when either the field ($H_z$) or the structure ($\epsilon^{-1}$)
    variable is considered separately Eq.~\ref{eq:act} has only one minimum
    (it is convex), and can be easily solved using standard methods.
We employ such an alternating directions strategy,
    where both $H_z$ and $\epsilon^{-1}$ are solved independently.
This process is extremely inefficient,
    but is employed because the underlying numerical methods
    do not require any tuning.
We expect considerable improvements in computational efficiency
    when more sophisticated algorithms are applied,
    especially those which can update $H_z$ and $\epsilon^{-1}$ independently.

@@

\section{Results}
The method described above is applied to 
    the design of three different waveguide couplers
    as shown in Figs.~\ref{fig:fiber}, \ref{fig:mode} and \ref{fig:aircore}.

% For all designs mention
% fast - 20 minutes
All results were obtained in 20 minutes on a single-core personal computer
% subwavelength? all use only 1-4 square vacuum wavelengths
% high performance

% Special mention (arbitrariness)
% breaking symmetry for figs 2, and 3
% fig 3 is not even index guided

% Starting point given just by e = 9
% Mention thresholding efficiencies.

% Coupling efficiency
% Design time
% Size of the device
% Frequency
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{quickfig/fiber} 
    \caption{Waveguide coupler for a wide, low-index 
        ($\epsilon=2.25$) waveguide. 
        The dielectric structure of the coupler and surrounding input and
        output waveguides is shown on the left, while the simulation
        validating our results is shown on the right.
        The coupler converts $96.3\%$ of the input power to the
        designated output mode.
        The device is extremely compact, 
        convering only $36 \times 66$ grid points,
        where the vacuum wavelength is 42 grid points.
        Computation time was 20 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{quickfig/mode-conv} 
    \caption{Coupler that converts the fundamental waveguide mode to the
        second-order waveguide mode.
        This problem is quite difficult since the two modes are of 
        opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
        (which has the same dimensions and vacuum wavelength as Fig. 1) 
        which achieves a coupling efficiency of $95.5\%$. 
        Computation time was 20 minutes on a personal computer. (movie?)
        }
    \label{fig:mode}
    \end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{quickfig/air-core}
    \caption{Coupler to an air-core mode.
        Here, not only are the modes of opposite symmetry,
        but the output waveguide operates on a fundamentally different
        principle (guided by Bragg reflection) than the input waveguide 
        (index guided).
        The device still achieves an efficiency of $83.3\%$, demonstrating the
        versatility of our method.
        The vacuum wavelength is 25 grid points, 
        while the device footprint is still $36 \times 66$ grid points.
        Computation time was 20 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}

>>>>

\documentclass[letterpaper,10pt]{article}
\usepackage{graphicx}
\usepackage{osameet2}
\usepackage{amsmath,amssymb}

\begin{document}
\title{Automated Design of Nanophotonic Waveguide Couplers}
\author{Jesse Lu and Jelena Vu\v{c}kovi\'{c}}
\address{Stanford University, Stanford, California, USA.}
\email{jesselu@stanford.edu}

\maketitle
\begin{abstract}
We introduce an ``objective-first'' approach to the design of
    nanophotonic waveguide couplers.
We show that our method is
    computationally fast (20 minutes), 
    requires no trial-and-error,
    does not require a good starting design,
    can be applied to arbitrary waveguide modes,
    and generates devices with high coupling efficiencies (typically 95\%)
    and small footprints (1-4 square vacuum wavelengths).
\end{abstract}
% \ocis{230.7370, 130.3990.}

@@

% Why is waveguide mode conversion important?
\subsection{The importance of waveguide mode conversion}
Optical mode conversion, 
    the efficient transfer of photons from one guided mode to another,
    is a fundamental requirement in nanophotonics.
Efficient conversion between waveguides modes
    is critical in the many cases, including:
    \begin{enumerate}
        \item Coupling to and from optical fiber\cite{}, 
        to communicate with the outside world.
    \item Coupling between various nanophotonic waveguides, 
        since different waveguides are best suited for different applications.
        For example, ridge waveguides seem ideal for low-loss transport\cite{},
            but other waveguides, 
            such as photonic crystal waveguides or slot waveguides,
            may be better suited for slow-light\cite{} 
            or energy-focusing devices\cite{}.
    \item Coupling between different materials such as
       passive, active\cite{}, and non-linear\cite{} materials. 
    \end{enumerate}
Although efficient waveguide coupling is essential in any nanophotonic system,
    practical methods to design couplers have not yet been developed.

% Why are current solutions insufficient?
\subsection{Approaches to designing waveguide couplers}
% Brute force
Of the design strategies currently available,
    brute-force parameter search is the most often employed 
    to-date because of its sheer simplicity\cite{}.
Although it may be suitable for tuning existing designs\cite{},
    the parameter space for most practical devices is simply too large
    for such a strategy to be tractable.

% Adiabatic mode conversion (large devices, symmetry breaking)
Adiabatic mode conversion strategies have been succesful
    for certain fiber-waveguide\cite{} and waveguide-waveguide\cite{} couplers,
    even if the resulting devices require large footprints.
However, adiabatic approaches cannot be used in many important cases 
    where the mode structure is more complex, such as
    coupling from ridge to some photonic crystal waveguides\cite{},
    or where the geometry of the problem is non-trivial, such as
    coupling in the out-of-plane direction, or
    coupling between modes of opposite symmetry.

% Local optimization
On the other hand, optimization methods based on local topological derivatives 
    seem very promising\cite{}
    in that they are both much faster than brute-force methods and
    more adaptable than adiabatic strategies.
However, these methods still carry a significant computational burden 
    in that every updated design must be simulated at least once.
Also, there is a significant burden for the user from whom
    a good initial design is usually required.

% How does your algorithm solve this problem?
% \subsection{Advantages of objective-first design}
In contrast, our method
\begin{itemize}
    \item does not employ brute-force parameter searches,
    \item does not require a good initial design,
    \item is computationally fast (no simulations required),
    \item generates couplers between seemingly arbitrary waveguide modes, and
    \item can generate these couplers in a very small footprint. 
\end{itemize}
We outline the general design strategy,
    as well as how it is applied to nanophotonic design, in the section below.

@@

The differences between Eqs.~\ref{eq:adj} and \ref{eq:ob1} are:
\begin{enumerate}
    \item in Eq.~\ref{eq:adj}, we attempt to decrease the design objective 
            (Eq.~\ref{eq:adj:obj}), 
        while in Eq.~\ref{eq:ob1}, the design objective is kept at zero 
            (Eq.~\ref{eq:ob1:con}); 
    \item in Eq.~\ref{eq:adj}, we always satisfy the underlying physics 
            (Eq.~\ref{eq:adj:con}),
        while in Eq.~\ref{eq:ob1}, physics is not satisfied,
            since the physics residual is generally non-zero
            (Eq.~\ref{eq:ob1:obj}).
\end{enumerate}

@@

The first practical implication of the objective-first approach 
    is that the number of independent variables is increased to include
    both $x$ and $p$.
In Eq.~\ref{eq:adj}, the constraint that physics must be satisfied, $g(x,p)=0$, 
    essentially forces $x$ to be a dependentent variable,
    since the choice of $p$ implicitly determines the value of $x$
    (there is generally a one-to-one mapping from $p$ to $x$).
In contrast, Eq.~\ref{eq:ob1} allows both $x$ and $p$ to vary independently,
    because the constraint, $f(x)=0$, is only a function of $x$.

@@

In conclusion, an objective-first approach results in a method 
    that is computationally fast (since it does not require simulation), and
    that can be applied to non-intuitive problems 
    where a functional starting design is not readily available.

% How is it applied to waveguide coupler design?
\subsection{Objective-first approach to waveguide coupler design}
We now apply an objective-first approach to the problem of designing 
    two-dimensional nanophotonic waveguide couplers.

@@

This problem contains many local minima (it is non-convex\cite{});
    however, when either the field ($H_z$) or the structure ($\epsilon^{-1}$)
    variable is considered separately Eq.~\ref{eq:act} has only one minimum
    (it is convex), and can be easily solved using standard methods
    such as employed in our previous work\cite{}.
We employ such an alternating directions strategy,
    where both $H_z$ and $\epsilon^{-1}$ are solved independently.
This process is extremely inefficient,
    but is employed because the underlying numerical methods
    do not require any tuning.
We expect considerable improvements in computational efficiency
    when more sophisticated algorithms are applied,
    especially those which can update $H_z$ and $\epsilon^{-1}$ independently.

@@

\section{Results}
The method described above is applied to 
    the design of three different waveguide couplers
    as shown in Figs.~\ref{fig:fiber}, \ref{fig:mode} and \ref{fig:aircore}.
These results demonstrate that our method
    \begin{itemize}
    \item is computationally fast, 
        requiring only 20 minutes on a single-core personal computer;
    \item generates compact devices with footprints of only
        1-4 square vacuum wavelengths;
    \item generates highly efficient devices with  
        typical coupling efficiencies of $95\%$;
    \item can be applied to arbitrary input and output modes,
        as seen from the diverse selection of desired output waveguide modes;
    \item does not require a good initial design.
        The initial design was simply $\epsilon = 9$ everywhere
            (a somewhat arbitrary guess, other values work as well),
            which in the case of Figs.~\ref{fig:mode} and \ref{fig:aircore}
            produces an initially non-functional device 
            (coupling efficiency of $0\%$).

    \end{itemize}

Lastly, note that the couplers in Figs.~\ref{fig:mode} and \ref{fig:aircore}
    must convert between modes of opposing symmetry,
    in which case adiabatic approaches cannot be applied.
Note also that the output waveguide in Fig.~\ref{fig:aircore}
    even operates on a completely different principle 
    (distributed reflection instead of index-guiding).

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{quickfig/fiber} 
    \caption{Waveguide coupler for a wide, low-index 
        ($\epsilon=2.25$) waveguide. 
        The dielectric structure of the coupler and surrounding input and
        output waveguides is shown on the left, while the simulation
        validating our results is shown on the right.
        The coupler converts $96.3\%$ of the input power to the
        designated output mode.
        The device is extremely compact, 
        convering only $36 \times 66$ grid points,
        where the vacuum wavelength is 42 grid points.
        Computation time was 20 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{quickfig/mode-conv} 
    \caption{Coupler that converts the fundamental waveguide mode to the
        second-order waveguide mode.
        This problem is quite difficult since the two modes are of 
        opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
        (which has the same dimensions and vacuum wavelength as Fig. 1) 
        which achieves a coupling efficiency of $95.5\%$. 
        Computation time was 20 minutes on a personal computer. (movie?)
        }
    \label{fig:mode}
    \end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{quickfig/air-core}
    \caption{Coupler to an air-core mode.
        Here, not only are the modes of opposite symmetry,
        but the output waveguide operates on a fundamentally different
        principle (guided by Bragg reflection) than the input waveguide 
        (index guided).
        The device still achieves an efficiency of $83.3\%$, demonstrating the
        versatility of our method.
        The vacuum wavelength is 25 grid points, 
        while the device footprint is still $36 \times 66$ grid points.
        Computation time was 20 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}

% \section{Conclusion}
% We introduce an objective-first approach to structure design,
%     and apply it to the design of nanophotonic waveguide couplers.
% We show that our method is
%     computationally fast,
%     generates high-efficiency devices with compact footprints,
%     does not require a good starting design, and
%     can be applied to arbitrary waveguide modes.
% 
%
